
We developed a benchmark suite containing various Rust language features
to test the SMACK extensions we developed.
\cref{tab:benchmarks} summarizes our benchmark suite.
Every category includes both correct and buggy benchmark versions.
Some notable included features are as follows:
\begin{itemize}
\item The closures category creates a closure that is passes to
  another function.
\item The generics category implements a generic trait for two
  generic structures.  A statically dispatched function is then
  invoked on the structures.
\item The vector category tests dynamic resizing and indexing
  of the Rust vector.
\item The cross-language category contains Rust
  programs that invoke C functions, including the \Cref{fig:crossfib} benchmark.
\item The ifc category contains the \textit{information
  flow control} (IFC) example from related \\ work~\cite{rust-safety17}. IFC models an
  access control method where access authority can only be
  increased. Using nondeterministic access levels, we verify that the
  IFC Rust implementation only allows access to the appropriate
  authority.
\end{itemize}
%
Currently, SMACK verifies most benchmarks in under 20 minutes.
The only exception is the full-blown IFC benchmark version that takes
several hours to complete.
%
The development of the benchmark suite helped us to identify
key language features that SMACK struggled with, and hence it guided our
efforts.
%
We plan to make our benchmark suite publicly available to bootstrap
the development of Rust verifiers.

